Palmer amaranth (\emph{Amaranthus palmeri}) adaptation to agroecossystems

### 2021-07-30 Manuscript/Manuscript.Rmd
fgfgf

Palmer amaranth (*Amaranthus palmeri* S. Watson) is currently considered one of the most economically damaged weed species to cropping systems in the United States [@ward2013]. The species has showed a remarkable capacity to evolve resistance to herbicides. Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017]. Uncontrolled Palmer amaranth in competition for water, light and nutrients can drastically reduce crop yields [@berger2015]. Palmer amaranth is documented with potential to reduce 91%, 68%, and 54% of corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields, respectively. Thus, Palmer amaranth poses a risk to sustainable agriculture.

Palmer amaranth is a fast growing summer annual forb indigenous to Sonoran Desert [@sauer1957]. The species would eventually emerge as a threat to US agriculture in the 1990s. Palmer amaranth weediness is likely a result of human-assisted selection in combination with species biology. Farm mechanization, conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems. On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@ward2013]. With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@oliveira2018]. Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017], and plant propagules [@yu2021]. Therefore, Palmer amaranth dispersal capacity make the species one of the most successful cases of weed adaption to cropping systems.

@@

Palmer amaranth is already documented in agronomic crops of South America [@larran2017; @kupper2017] and Southern Europe [@milani2021]. In the US, Palmer amaranth is established at crop [@garetson2019] and non-crop land [@bagavathiannan2016] in the warm southern United States but its range is expanding to cool temperatures northward. For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a],  Michigan [@kohrt2017], and Connecticut [@aulakh2021]. Successful cases of Palmer amaranth invasion and near to eradication is reported in Minnesota [@yu2021]. No Palmer amaranth actively growing was found in Canada; however, Palmer amaranth seeds was detected in sweet potato slips [@page2021]. Nonetheless, it seems fated the need to manage Palmer amaranth in agronomic crops throughout multiple environments in the near future. Therefore, strategies on Palmer amaranth management should encompass the agroecosystem level but not only attempts to eradicate the weed. Most tactics to manage Palmer amaranth are based on technology fixes [@scott2011], which are short-term (e.g., herbicide and/or tillage) rather than long-term weed management.

@@

Palmer amaranth cumulative flowering from first transplant timing ranged from late June to early September (Figure 2). Palmer amaranth growing in corn resulted in a longer flowering pattern compared to fallow and soybean. Palmer amaranth reached 10% flowering in soybean at doy 180, which was slightly different from fallow (doy 180.9; *P* = 0.01) and corn (doy 181.7; *P* = 0.00). Nonetheless, the 10% cumulative Palmer amaranth flowering in soybean, fallow and corn occurred at the end of June. The 50% Palmer amaranth cumulative flowering occurred in July. For example, Palmer amaranth reached 50% in fallow at doy 193.4, followed by soybean (doy 194.8), corn (doy 206.6). Similar trend was observed at 90% Palmer amaranth cumulative flowering. Palmer amaranth growing in corn reached 90% flowering at doy 252.6 (early September), which was 37.8 and 32.2 days after Palmer amaranth 90% flowering in fallow and soybean, respectively.

@@

Palmer amaranth accumulated more biomass when growing in fallow compared to Palmer amaranth growing in soybean and corn. At first transplanting time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g plant-1 and 14.9 g plant-1 in fallow, soybean and corn, respectively. Palmer amaranth produce less biomass at second transplanting time. For example, Palmer amaranth growing in fallow resulted in 49.2 g plant-1, 3.4 g plant-1 in soybean and 1.4 g plant when growing in corn.

@@

>>>>

Palmer amaranth (*Amaranthus palmeri* S. Watson) is currently considered one of the most economically damaged weed species to cropping systems in the United States [@ward2013]. The species has showed a remarkable capacity to evolve resistance to herbicides. Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017]. Uncontrolled Palmer amaranth in competition for water, light and nutrients can drastically impact on crop yields [@berger2015]. Palmer amaranth is documented with potential to reduce 91%, 68%, and 54% of corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields, respectively. Thus, unmanaged Palmer amaranth poses an economical risk to sustainable agriculture.

Palmer amaranth is a fast growing summer annual forb indigenous to Sonoran Desert [@sauer1957]. The species would eventually emerge as a threat to US agriculture in the 1990s. Palmer amaranth weediness is likely a result of human-assisted selection in combination with species biology. Farm mechanization, conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems. On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@ward2013]. With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@oliveira2018]. Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017], and plant propagules [@yu2021]. Palmer amaranth dispersal capacity make the species one of the most successful cases of weed adaption to cropping systems.

@@

Palmer amaranth is already documented in agronomic crops of South America [@larran2017; @kupper2017] and Southern Europe [@milani2021]. In the US, Palmer amaranth is well established at crop [@garetson2019] and non-crop land [@bagavathiannan2016] in the warm southern United States but its range is expanding to cool temperatures northward. For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a],  Michigan [@kohrt2017], and Connecticut [@aulakh2021]. Successful cases of Palmer amaranth invasion and near to eradication is reported in Minnesota [@yu2021]. No Palmer amaranth actively growing was found in Canada; however, Palmer amaranth seeds was detected in sweet potato slips [@page2021]. Nonetheless, it seems fated the need to manage Palmer amaranth in agronomic crops throughout multiple environments in the near future. Therefore, strategies on Palmer amaranth management should encompass the agroecosystem level but not only attempts to eradicate the weed. Most tactics to manage Palmer amaranth are based on technology fixes [@scott2011], which are short-term (e.g., herbicide and/or tillage) rather than long-term weed management. Long-term tactics to minimize the negative Palmer amaranth impact in the agroecossystem should include species strategies to adapt and grow into different environments.

@@

Palmer amaranth growing in corn resulted in a longer flowering pattern compared to fallow and soybean (Figure 2A). Palmer amaranth reached 10% flowering in soybean at doy 180, which was slightly different from fallow (doy 180.9; *P* = 0.01) and corn (doy 181.7; *P* = 0.00). Nonetheless, the 10% cumulative Palmer amaranth flowering in soybean, fallow and corn occurred at the end of June. The 50% Palmer amaranth cumulative flowering occurred in July. For example, Palmer amaranth reached 50% in fallow at doy 193.4, followed by soybean (doy 194.8), corn (doy 206.6). Similar trend was observed at 90% Palmer amaranth cumulative flowering. Palmer amaranth growing in corn reached 90% flowering at doy 252.6 (early September), which was 37.8 and 32.2 days after Palmer amaranth 90% flowering in fallow and soybean, respectively.

@@

Palmer amaranth accumulated more biomass when growing in fallow compared to Palmer amaranth growing in soybean and corn (Figure 3). At first transplanting time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g plant-1 and 14.9 g plant-1 in fallow, soybean and corn, respectively. Palmer amaranth produce less biomass at second transplanting time. For example, Palmer amaranth growing in fallow resulted in 49.2 g plant-1, 3.4 g plant-1 in soybean and 1.4 g plant when growing in corn.

@@

Our study showed the Palmer amaranth adaptation plasticity when growing into different environments. In general,  These result highlight the plant response to grow degree days.

Palmer amaranth growing plasticity varied within time and space. In general, Palmer amaranth produced more biomass and taller plants when transplanted in early June rather than July. In addition, at first cohort, when resources (e.g., soil nutrients) and conditions (e.g., light) were not limited for both species, Palmer amaranth growing strategy was likely to mimic the cropping systems (Figure 4). In contrast, at second cohort, Palmer amaranth grow and development was limited due to the competitive advantages from the crop, including closed canopy. For example, when growing in corn at first cohort, Palmer amaranth plants accumulated lowest biomass, but produce tallest plants. Higher Palmer amaranth plants in corn is likely the species strategy to compete for light in between corn rows. At first cohort, corn and Palmer amaranth had abundant light over the canopy to grow and develop; however, at second cohort, Palmer amaranth was transplanted when corn canopy was nearly closed, which reduced Palmer amaranth competition. As a result, Palmer amaranth height and biomass was nearly 50% lower compared to its first cohort.

Similar trend was observed to Palmer amaranth growing in soybeans. The difference in Palmer amaranth height and biomass was lower due to soybean growing strategy.

Palmer amaranth produced the highest biomass when growing without crop competition (fallow). When growing in fallow, cohort time had low impact on Palmer amaranth growing and development. Moreover, Palmer amaranth biomass was more variable when growing in fallow.

Palmer amaranth cumulative flowering from first transplant timing ranged from late June to early September.

### 2021-07-30 Manuscript/Manuscript.tex
fgfgf

Palmer amaranth (\emph{Amaranthus palmeri} S. Watson) is currently
considered one of the most economically damaged weed species to cropping
systems in the United States (Ward et al., 2013). The species has showed
a remarkable capacity to evolve resistance to herbicides. Palmer
amaranth has evolved resistance to eight herbicide sites of action
(Heap, 2021), increasing the weed management complexity (Lindsay et al.,
2017). Uncontrolled Palmer amaranth in competition for water, light and
nutrients can drastically reduce crop yields (Berger et al., 2015).
Palmer amaranth is documented with potential to reduce 91\%, 68\%, and
54\% of corn (Massinga et al., 2001), soybean (Klingaman and Oliver,
1994), and cotton (Morgan et al., 2001) yields, respectively. Thus,
Palmer amaranth poses a risk to sustainable agriculture.

Palmer amaranth is a fast growing summer annual forb indigenous to
Sonoran Desert (Sauer, 1957). The species would eventually emerge as a
threat to US agriculture in the 1990s. Palmer amaranth weediness is
likely a result of human-assisted selection in combination with species
biology. Farm mechanization, conservation agriculture (e.g., no-till),
and reliance on herbicides for weed management are the main
human-mediated selection of Palmer amaranth into cropping systems. On
the other hand, Palmer amaranth is a prolific seed producer with a C4
photosynthetic apparatus (Ward et al., 2013). With a dioecy nature,
Palmer amaranth male and female plants are obligate outcrosser species,
increasing the chances of exchanging adaptive traits among plants
(Oliveira et al., 2018). Also, Palmer amaranth small seed size (e.g, 1
mm) tend to thrive in no-tillage systems (Price et al., 2011), and
spread across locations through farm equipment (Sauer, 1972), manure
(Hartzler and Anderson, 2016), animals (Farmer et al., 2017), and plant
propagules (Yu et al., 2021). Therefore, Palmer amaranth dispersal
capacity make the species one of the most successful cases of weed
adaption to cropping systems.

@@

Palmer amaranth is already documented in agronomic crops of South
America (Larran et al., 2017; Küpper et al., 2017) and Southern Europe
(Milani et al., 2021). In the US, Palmer amaranth is established at crop
(Garetson et al., 2019) and non-crop land (Bagavathiannan and
Norsworthy, 2016) in the warm southern United States but its range is
expanding to cool temperatures northward. For example, herbicide
resistant Palmer amaranth is widespread in Nebraska (Oliveira et al.,
2021), Michigan (Kohrt et al., 2017), and Connecticut (Aulakh et al.,
2021). Successful cases of Palmer amaranth invasion and near to
eradication is reported in Minnesota (Yu et al., 2021). No Palmer
amaranth actively growing was found in Canada; however, Palmer amaranth
seeds was detected in sweet potato slips (Page et al., 2021).
Nonetheless, it seems fated the need to manage Palmer amaranth in
agronomic crops throughout multiple environments in the near future.
Therefore, strategies on Palmer amaranth management should encompass the
agroecosystem level but not only attempts to eradicate the weed. Most
tactics to manage Palmer amaranth are based on technology fixes (Scott,
2011), which are short-term (e.g., herbicide and/or tillage) rather than
long-term weed management.

@@

Monthly mean air temperature and sum precipitation were obtained using
Daymet weather data from June through September across the five
locations in 2018 and 2019 (\textbf{correndo2021?}) (Figure 1)

@@

Palmer amaranth cumulative flowering from first transplant timing ranged
from late June to early September (Figure 2). Palmer amaranth growing in
corn resulted in a longer flowering pattern compared to fallow and
soybean. Palmer amaranth reached 10\% flowering in soybean at doy 180,
which was slightly different from fallow (doy 180.9; \emph{P} = 0.01)
and corn (doy 181.7; \emph{P} = 0.00). Nonetheless, the 10\% cumulative
Palmer amaranth flowering in soybean, fallow and corn occurred at the
end of June. The 50\% Palmer amaranth cumulative flowering occurred in
July. For example, Palmer amaranth reached 50\% in fallow at doy 193.4,
followed by soybean (doy 194.8), corn (doy 206.6). Similar trend was
observed at 90\% Palmer amaranth cumulative flowering. Palmer amaranth
growing in corn reached 90\% flowering at doy 252.6 (early September),
which was 37.8 and 32.2 days after Palmer amaranth 90\% flowering in
fallow and soybean, respectively.

@@

@@

>>>>

Palmer amaranth (\emph{Amaranthus palmeri} S. Watson) is currently
considered one of the most economically damaged weed species to cropping
systems in the United States (Ward et al., 2013). The species has showed
a remarkable capacity to evolve resistance to herbicides. Palmer
amaranth has evolved resistance to eight herbicide sites of action
(Heap, 2021), increasing the weed management complexity (Lindsay et al.,
2017). Uncontrolled Palmer amaranth in competition for water, light and
nutrients can drastically impact on crop yields (Berger et al., 2015).
Palmer amaranth is documented with potential to reduce 91\%, 68\%, and
54\% of corn (Massinga et al., 2001), soybean (Klingaman and Oliver,
1994), and cotton (Morgan et al., 2001) yields, respectively. Thus,
unmanaged Palmer amaranth poses an economical risk to sustainable
agriculture.

Palmer amaranth is a fast growing summer annual forb indigenous to
Sonoran Desert (Sauer, 1957). The species would eventually emerge as a
threat to US agriculture in the 1990s. Palmer amaranth weediness is
likely a result of human-assisted selection in combination with species
biology. Farm mechanization, conservation agriculture (e.g., no-till),
and reliance on herbicides for weed management are the main
human-mediated selection of Palmer amaranth into cropping systems. On
the other hand, Palmer amaranth is a prolific seed producer with a C4
photosynthetic apparatus (Ward et al., 2013). With a dioecy nature,
Palmer amaranth male and female plants are obligate outcrosser species,
increasing the chances of exchanging adaptive traits among plants
(Oliveira et al., 2018). Also, Palmer amaranth small seed size (e.g, 1
mm) tend to thrive in no-tillage systems (Price et al., 2011), and
spread across locations through farm equipment (Sauer, 1972), manure
(Hartzler and Anderson, 2016), animals (Farmer et al., 2017), and plant
propagules (Yu et al., 2021). Palmer amaranth dispersal capacity make
the species one of the most successful cases of weed adaption to
cropping systems.

@@

Palmer amaranth is already documented in agronomic crops of South
America (Larran et al., 2017; Küpper et al., 2017) and Southern Europe
(Milani et al., 2021). In the US, Palmer amaranth is well established at
crop (Garetson et al., 2019) and non-crop land (Bagavathiannan and
Norsworthy, 2016) in the warm southern United States but its range is
expanding to cool temperatures northward. For example, herbicide
resistant Palmer amaranth is widespread in Nebraska (Oliveira et al.,
2021), Michigan (Kohrt et al., 2017), and Connecticut (Aulakh et al.,
2021). Successful cases of Palmer amaranth invasion and near to
eradication is reported in Minnesota (Yu et al., 2021). No Palmer
amaranth actively growing was found in Canada; however, Palmer amaranth
seeds was detected in sweet potato slips (Page et al., 2021).
Nonetheless, it seems fated the need to manage Palmer amaranth in
agronomic crops throughout multiple environments in the near future.
Therefore, strategies on Palmer amaranth management should encompass the
agroecosystem level but not only attempts to eradicate the weed. Most
tactics to manage Palmer amaranth are based on technology fixes (Scott,
2011), which are short-term (e.g., herbicide and/or tillage) rather than
long-term weed management. Long-term tactics to minimize the negative
Palmer amaranth impact in the agroecossystem should include species
strategies to adapt and grow into different environments.

@@

Monthly mean air temperature and sum precipitation were obtained using
Daymet weather data from June through September across the five
locations in 2018 and 2019 (Correndo et al., 2021) (Figure 1)

@@

Palmer amaranth growing in corn resulted in a longer flowering pattern
compared to fallow and soybean (Figure 2A). Palmer amaranth reached 10\%
flowering in soybean at doy 180, which was slightly different from
fallow (doy 180.9; \emph{P} = 0.01) and corn (doy 181.7; \emph{P} =
0.00). Nonetheless, the 10\% cumulative Palmer amaranth flowering in
soybean, fallow and corn occurred at the end of June. The 50\% Palmer
amaranth cumulative flowering occurred in July. For example, Palmer
amaranth reached 50\% in fallow at doy 193.4, followed by soybean (doy
194.8), corn (doy 206.6). Similar trend was observed at 90\% Palmer
amaranth cumulative flowering. Palmer amaranth growing in corn reached
90\% flowering at doy 252.6 (early September), which was 37.8 and 32.2
days after Palmer amaranth 90\% flowering in fallow and soybean,
respectively.

@@

Palmer amaranth cumulative flowering from first transplant timing ranged
from late June to early September.

@@

\leavevmode\hypertarget{ref-correndo2021}{}%
Correndo, A. A., Moro Rosso, L. H., and Ciampitti, I. A. (2021).
Retrieving and processing agro-meteorological data from {API}-client
sources using {R} software. \emph{BMC Research Notes} 14, 205.
doi:\href{https://doi.org/10.1186/s13104-021-05622-8}{10.1186/s13104-021-05622-8}.
